### 2022-11-24 dissertation/chapters/06-bayesian-hyper-heuristic.tex
✨ (final-review): Up to chapter 4

\begin{quotation}
      \noindent ``The result is a posterior distribution, which the agent may use as its new prior in the next step.''
\end{quotation}
\begin{flushright}
      - Pedro Domingos, The Master Algorithm
\end{flushright}

\noindent
The above quote was the inspiration for the development of a novel \acf{HH} that uses Bayesian probability concepts as a selection mechanism to drive the heuristic selection process. Thus far the reader has been presented with all of the necessary background information on \acp{ANN} (Chapter \ref{chap:anns}), low-level \index{heuristic}heuristics (Chapter \ref{chap:heuristics}), \acp{HH} (Chapter \ref{chap:hhs}) and lastly, probability theory (Chapter \ref{chap:probability}). These elements form the fundamental components of the proposed \Acf{BHH}. This chapter provides the details around the implementation of the \Ac{BHH} and explains how it used to train \acp{FFNN}. The remainder of the chapter is structured as follows.

@@

Formal classification of the \Acs{BHH} is needed. Chapter \ref{chap:hhs} presented the reader with a proposed classification scheme for \acp{HH} by~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010}. According to the aforementioned classification scheme, the \Acs{BHH} is a population-based, meta-hyper-heuristic that utilises selection and perturbation of low-level heuristics in an online learning fashion. A breakdown of the classification is given as follows.

@@

\noindent
The following sections presents the reader with more detail on the implementation of the \acs{BHH}. Discussions lead with the \acs{BHH} architecture and general \acs{HH} framework in the following section.

@@

\noindent
With reference to Figure \ref{fig:bhh_architecture}, the components are briefly given below in the order of information flow during the \acs{FFNN} training process steps.

@@

\noindent
Each of the components in Figure \ref{fig:bhh_architecture} is discussed in detail in the following sections.

@@

\noindent
Primary and proxied state parameters must be maintained together. Since entities represent candidate solutions, which are forms of state, entities are ideal candidates to store and maintain these state parameters. Entities are extended to include the primary state elements of \textit{all} the underlying low-level heuristics, as well as the candidate solution (weights) for the \acs{FFNN}. At each heuristic-entity application step, all state parameters, per entity, are updated either by primary method or by proxied method. The \acs{BHH} thus incorporates a mapping of proxied state update operation as given in the example in Table \ref{tab:bhh:heuristic_pool:proxy_mapping_example} below.

@@

\noindent
From the example given in Table \ref{tab:bhh:heuristic_pool:proxy_mapping_example}, when \index{heuristic}heuristic A is selected, it will outsource state update operations from \index{heuristic}heuristic B, for state parameter 2. Heuristic B will outsource from \index{heuristic}heuristic A, for state parameter 3. Finally, \index{heuristic}heuristic C will outsource from \index{heuristic}heuristic A and B, for state parameters 2 and 3 respectively. In this way, all \index{heuristic}heuristics maintain all the state parameters.

@@

\noindent
From the list above, it should be clear that entity state becomes increasingly complicated with the increase of the number of distinct \index{heuristic}heuristics in the \index{heuristic pool}heuristic pool with unique state parameters.

@@

\noindent
Entities in the entity pool are each assigned an associated \index{heuristic}heuristic. The \acs{BHH} selects from the \index{heuristic pool}heuristic pool a low-level \index{heuristic}heuristic to be applied to an individual entity. The outcome of this selection process is a mapping table that tracks which \index{heuristic}heuristic has been selected for which entity. The selection process is executed by the selection mechanism of the \acs{BHH}. These \index{heuristic}heuristic-entity combinations are applied to the underlying \acs{FFNN}. The \acs{BHH} tracks the performance of each of these combinations throughout the training process in a performance log. The performance log is presented in the following section.

@@

\noindent
An example of the performance log implemented by the \acs{BHH} is given in Table \ref{tab:bhh:performance_log:performance_log_example} below.

@@

\noindent
The performance log itself introduces a number of design considerations as well. The larger the performance log, the longer memory is retained throughout the training process. The performance log size, referred to as the \textit{replay window size}, controls the recency of evidence from which the \acs{BHH} must learn. If the replay window size is too small, it does not accumulate enough samples/evidence to statistically make accurate selections. If the replay window size is too big, a bias could exist for selecting \index{heuristic}heuristics that performed well in the past. Furthermore, it is not guaranteed that past performance is indicative of future performance during the training process. The performance log should be considered along with the selection of a credit assignment strategy. Credit assignment strategies are presented in the following section.

@@

\noindent
The implementation of a credit assignment strategy is simply a function that translates the performance log from Table \ref{tab:bhh:performance_log:performance_log_example}, into Table \ref{tab:bhh:credit_assignment_strategy:credit_assignment_example}.

@@

\noindent
Credit is used by the selection mechanism of the \acs{BHH}. The selection mechanism is presented in the following section.

@@

\noindent
It should be noted that event information is stored and observed directly from the performance log as presented in Table \ref{tab:bhh:credit_assignment_strategy:credit_assignment_example}. From the above list of events, the event $C$ is dependent on the occurrence of $H$ and $E$ and a credit assignment strategy.

@@

\noindent
Since the \acs{BHH} implements a Bayesian classifier, the \acs{BHH} is not concerned with actual probabilities. Equation~\eqref{eq:bhh:selection_mechanism:bayes_theorem} can thus be evaluated for proportionality. The resulting proportionality is expressed in Equation~\eqref{eq:bhh:selection_mechanism:bayes_theorem_prop_to} below.

@@

\noindent
The conditionality implemented by \index{Bayes' Theorem}Bayes' Theorem can be extended to include multiple criteria. \index{Bayes' Theorem}Bayes' Theorem is used by the selection mechanism of the \acs{BHH}. The predictive model that is used by the selection mechanism of the \acs{BHH} is presented in the following section.

@@

This section presents the predictive model as implemented by the selection mechanism of the \acs{BHH}. The predictive model is arguably the most important component of the \acs{BHH}, as it drives the heuristic selection process. The \acs{BHH} implements a predictive model that predicts which heuristic to select, given the conditionality that a particular entity is used and a particular credit assignment criteria is met. The predictive model can be derived as follows.\\\\
\textbf{Setup}:
\\
\noindent
Let

@@

\noindent
Vector notation is dropped for sake of convenience, simplicity and readability. The parameterised predictive model, as derived from \index{Bayes' Theorem}Bayes' theorem, is then given in Equation~\eqref{eq:bhh:selection_mechanism:predictive_model} below.

@@

\noindent
Notice the joint probability over events $E$ and $C$, given the selection of a heuristic $H$, denoted by the product of the sums of the separate parts for $E$ and $C$, in the denominator. The calculation in the denominator can be intractable. Since the \acs{BHH} selection mechanism uses the predictive model as a classifier, the posterior distribution as given in Equation~\eqref{eq:bhh:selection_mechanism:bayes_theorem}, can be evaluated for proportionality, as is given in Equation~\eqref{eq:bhh:selection_mechanism:predictive_model_prop_to} below.

@@

\noindent
The predictive model thus models the \textit{proportional} probability of the event (selection of) heuristic $H$, given allocation to entity $E$ and credit $C$, parameterised by sampled $\theta$, $\phi$ and $\psi$. These parameters are in turn parameterised by concentrations $\alpha$, $\beta$, $\gamma_{1}$ and $\gamma_{0}$, which denote the prior beliefs of the \acs{BHH}.

@@

\noindent
In Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:h_pmf} above, $N_{k}$ is a summary variable such that $N_{k} = \sum_{i=i}^{I} \mathbbm{1}_{1}(h_{i,k})$, denoting the count of occurrences of the event $h_{i}$ taking on class $k$ in $I$ independent, identical runs.

@@

\noindent
In Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:EgH_pmf} above, $N_{j,k}$ is a summary variable such that $N_{j,k} = \sum_{i=i}^{I} \mathbbm{1}_{1}(e_{i,j,k})\mathbbm{1}_{1}(h_{i,k})$, denoting the count of occurrences of the events $e_{i}$ taking on class $j$ and $h_{i}$ taking on class $k$, i.e. the count of occurrences of both entity $e$ and heuristic $k$ occurring together in $I$ independent, identical runs.

@@

\noindent
In Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:CgH_pmf} above, $N_{k}$ is the same summary variable as described for Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:h_pmf}. $N_{1,k}$ is a summary variable such that $N_{1,k} = \sum_{i=1}^{I} \mathbbm{1}_{1}(c_{i,k})\mathbbm{1}_{1}(h_{i,k})$, denoting the count of occurrences of the events $c_{i}$ taking on a success $(c_{i}=1)$ and $h_{i}$ taking on class $k$, i.e. the count of occurrences of both succeeding in the performance criteria and heuristic $k$ occurring together in $I$ independent, identical runs. Similarly $N_{0,k} = N_{k} - N_{1,k}$ would denote the count of occurrences of the events $c_{i}$ taking on a failure $(c_{i}=0)$ and $h_{i}$
taking on class $k$.

@@

\noindent
Consider the practical implementation of the predictive model as shown in Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:HgEC_pmf} above. Computationally the equation presented in Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:HgEC_pmf} will underflow on a real computer if the resulting probabilities are very small. Discussions follow on numerical stability as it relates to practical implementation.

@@

\noindent
The log-sum-exp trick as shown above caters for very small probabilities, however, there might be a situation where  a random event is never seen, purely by chance. This results in a scenario referred to as \textit{mode collapse} and is discussed next.

@@

\noindent
Another suggestion is to ensure that at least one of every heuristic is always selected during training. The \acs{BHH} does not incorporate this approach as this could result in a situation where a bad choice of heuristic is continually used throughout the training process, possibly resulting in bad update steps/selections. Instead, the \acs{BHH} relies on the underlying learning process to control the selective pressure according to performance and nothing else. If this leads to a situation where mode collapse occurs, the \acs{BHH} accepts the outcome.

@@

\noindent
The derivations of the \acs{MLE} for $\hat{\theta}_{k}$, $\hat{\phi}_{j,k}$ and
$\hat{\psi}_{k}$ are presented as follows. The log likelihood of $\hat{\theta}$, $\hat{\phi}$ and $\hat{\psi}$ as derived from the Equation~\eqref{eq:bhh:selection_mechanism:predictive_model_prop_to} is given in Equation~\eqref{eq:bhh:optimisation_step:mle:log_likelihood_all} below.

@@

\noindent
Equation~\eqref{eq:bhh:optimisation_step:mle:log_likelihood_all} can be broken down into each of its components. Consider the log likelihood of $\psi$ as denoted by Equation~\eqref{eq:bhh:optimisation_step:mle:log_likelihood_psi} below.

@@

\noindent
The \acs{MLE} for $\hat{\psi_{k}} $ can be calculated by taking the partial derivative of Equation~\eqref{eq:bhh:optimisation_step:mle:log_likelihood_psi} with respect to $\psi_{k}$ and equating to zero as presented in Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_psi} as follows.

@@

\noindent
The \acs{MLE} for $\hat{\theta_{k}} $ can be calculated similarly. However, one has to compensate for the $K-1$ simplex $S$ by adding error factor $\epsilon$, to correct values for $\theta$, where $\sum_{k}^{K} \theta_{k} \neq 1$. The new log likelihood function is then given in Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_theta_part_1} below.

@@

\noindent
Solving first for $\epsilon$ yields Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_theta_part_2} below.

@@

\noindent
Then solving for $\theta_{k}$ yields Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_theta_part_3} as presented below.

@@

\noindent
Substituting $N = \epsilon$ back into Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_theta_part_3} yields Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_theta_part_4} below.

@@

\noindent
The \acs{MLE} for $\hat{\phi}_{j, k} $ is calculated similarly. To compensate for the $K-1$ simplex $S$, an error $\lambda = (\lambda_{1}, \dots, \lambda_{J})$ is added as is shown in Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_phi_part_1} below.

@@

\noindent
Solving first for $\lambda_{j}$ yields Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_phi_part_2} as is given below.

@@

\noindent
Then solving for $\phi_{j,k}$ yields Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_phi_part_3} below.

@@

\noindent
Substituting $N_{j} = \lambda_{j}$ back into Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_phi_part_3} yields Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_phi_part_4} below.

@@

\noindent
The event $H$ is a multinomial distribution with probability parameter $\theta$. It is known that the conjugate prior to a multinomial distribution is a \index{Dirichlet probability distribution}Dirichlet probability distribution. The \textit{prior} probability distribution for $\theta_{k}$ is thus presented in Equation~\eqref{eq:bhh:optimisation_step:map:map_theta_prior} below.

@@

\noindent
Furthermore, the event $E$ is also a multinomial distribution with parameter $\phi$. The \textit{prior} probability distribution for $\phi_{j,k}$ is thus presented in Equation~\eqref{eq:bhh:optimisation_step:map:map_phi_prior} below.

@@

\noindent
Finally, the event $C$ is a \index{Binomial probability distribution}Binomial probability distribution with probability parameter $\psi$. It is known that the conjugate prior to a \index{Binomial probability distribution}Binomial probability distribution is a Beta probability distribution. The prior probability distribution for $\psi_{k}$ is thus presented in Equation~\eqref{eq:bhh:optimisation_step:map:map_psi_prior} below.

@@

\noindent
Putting the likelihood and priors together yields the posterior distribution as given in Equation~\eqref{eq:bhh:optimisation_step:map:map_psi_prior} below.

@@

\noindent
From Equation~\eqref{eq:bhh:optimisation_step:map:posterior} above, it can be seen that the posterior distribution has the same form $\mathcal{A}(v)$ as the prior distribution. The concentration update operations can then be given in Equations \eqref{eq:bhh:optimisation_step:map:alpha_update_operation}, \eqref{eq:bhh:optimisation_step:map:beta_update_operation}, \eqref{eq:bhh:optimisation_step:map:gamma1_update_operation} and \eqref{eq:bhh:optimisation_step:map:gamma2_update_operation} below.

@@

\noindent
It can be seen that the prior and posterior probability distributions are of the same form, $\mathcal{A}(v)$. It can therefore be said that the \acs{BHH} implements a Gaussian process. Since the reselection of heuristics happens at regular intervals, the outcome of a selection in one iteration may influence the outcome of another in the next iteration, making the implementation of the \acs{BHH} a \acf{HMM}.

@@

\noindent
The design choices at a high level is not as strict for the \acs{BHH} as it is for low-level heuristics. If one is unsure as to which low-level hyper-parameters to use, one could simply include multiple configurations of a low-level heuristic, each with its own unique hyper-parameter configurations.

@@

\noindent
A number of other hyper-parameters are specifically included to address the list of considerations as given above. These include \textit{reselection}, \textit{replay} and \textit{reanalysis window sizes}, \textit{burn in}, \textit{discounted rewards} and \textit{normalisation}. Each of these are presented next.

@@

\noindent
The following section discusses the default values for proxied update step operations.

>>>>

\begin{quotation}
      ``The result is a posterior distribution, which the agent may use as its new prior in the next step.''
\end{quotation}
\begin{flushright}
      - Pedro Domingos, The Master Algorithm
\end{flushright}

The above quote was the inspiration for the development of a novel \acf{HH} that uses Bayesian probability concepts as a selection mechanism to drive the heuristic selection process. Thus far the reader has been presented with all of the necessary background information on \acp{ANN} (Chapter \ref{chap:anns}), low-level \index{heuristic}heuristics (Chapter \ref{chap:heuristics}), \acp{HH} (Chapter \ref{chap:hhs}) and lastly, probability theory (Chapter \ref{chap:probability}). These elements form the fundamental components of the proposed \Acf{BHH}. This chapter provides the details around the implementation of the \Ac{BHH} and explains how it used to train \acp{FFNN}. The remainder of the chapter is structured as follows:

@@

Formal classification of the \Acs{BHH} is needed. Chapter \ref{chap:hhs} presented the reader with a proposed classification scheme for \acp{HH} by~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010}. According to the aforementioned classification scheme, the \Acs{BHH} is a population-based, meta-hyper-heuristic that utilises selection and perturbation of low-level heuristics in an online learning fashion. A breakdown of the classification is given as follows:

@@

The following sections presents the reader with more detail on the implementation of the \acs{BHH}. Discussions lead with the \acs{BHH} architecture and general \acs{HH} framework in the following section.

@@

With reference to Figure \ref{fig:bhh_architecture}, the components are briefly given below in the order of information flow during the \acs{FFNN} training process steps.

@@

Each of the components in Figure \ref{fig:bhh_architecture} is discussed in detail in the following sections.

@@

Primary and proxied state parameters must be maintained together. Since entities represent candidate solutions, which are forms of state, entities are ideal candidates to store and maintain these state parameters. Entities are extended to include the primary state elements of \textit{all} the underlying low-level heuristics, as well as the candidate solution (weights) for the \acs{FFNN}. At each heuristic-entity application step, all state parameters, per entity, are updated either by primary method or by proxied method. The \acs{BHH} thus incorporates a mapping of proxied state update operation as given in the example in Table \ref{tab:bhh:heuristic_pool:proxy_mapping_example} below.

@@

From the example given in Table \ref{tab:bhh:heuristic_pool:proxy_mapping_example}, when \index{heuristic}heuristic A is selected, it will outsource state update operations from \index{heuristic}heuristic B, for state parameter 2. Heuristic B will outsource from \index{heuristic}heuristic A, for state parameter 3. Finally, \index{heuristic}heuristic C will outsource from \index{heuristic}heuristic A and B, for state parameters 2 and 3 respectively. In this way, all \index{heuristic}heuristics maintain all the state parameters.

@@

From the list above, it should be clear that entity state becomes increasingly complicated with the increase of the number of distinct \index{heuristic}heuristics in the \index{heuristic pool}heuristic pool with unique state parameters.

@@

Entities in the entity pool are each assigned an associated \index{heuristic}heuristic. The \acs{BHH} selects from the \index{heuristic pool}heuristic pool a low-level \index{heuristic}heuristic to be applied to an individual entity. The outcome of this selection process is a mapping table that tracks which \index{heuristic}heuristic has been selected for which entity. The selection process is executed by the selection mechanism of the \acs{BHH}. These \index{heuristic}heuristic-entity combinations are applied to the underlying \acs{FFNN}. The \acs{BHH} tracks the performance of each of these combinations throughout the training process in a performance log. The performance log is presented in the following section.

@@

An example of the performance log implemented by the \acs{BHH} is given in Table \ref{tab:bhh:performance_log:performance_log_example} below.

@@

The performance log itself introduces a number of design considerations as well. The larger the performance log, the longer memory is retained throughout the training process. The performance log size, referred to as the \textit{replay window size}, controls the recency of evidence from which the \acs{BHH} must learn. If the replay window size is too small, it does not accumulate enough samples/evidence to statistically make accurate selections. If the replay window size is too big, a bias could exist for selecting \index{heuristic}heuristics that performed well in the past. Furthermore, it is not guaranteed that past performance is indicative of future performance during the training process. The performance log should be considered along with the selection of a credit assignment strategy. Credit assignment strategies are presented in the following section.

@@

The implementation of a credit assignment strategy is simply a function that translates the performance log from Table \ref{tab:bhh:performance_log:performance_log_example}, into Table \ref{tab:bhh:credit_assignment_strategy:credit_assignment_example}.

@@

Credit is used by the selection mechanism of the \acs{BHH}. The selection mechanism is presented in the following section.

@@

It should be noted that event information is stored and observed directly from the performance log as presented in Table \ref{tab:bhh:credit_assignment_strategy:credit_assignment_example}. From the above list of events, the event $C$ is dependent on the occurrence of $H$ and $E$ and a credit assignment strategy.

@@

Since the \acs{BHH} implements a Bayesian classifier, the \acs{BHH} is not concerned with actual probabilities. Equation~\eqref{eq:bhh:selection_mechanism:bayes_theorem} can thus be evaluated for proportionality. The resulting proportionality is expressed in Equation~\eqref{eq:bhh:selection_mechanism:bayes_theorem_prop_to} below.

@@

The conditionality implemented by \index{Bayes' Theorem}Bayes' Theorem can be extended to include multiple criteria. \index{Bayes' Theorem}Bayes' Theorem is used by the selection mechanism of the \acs{BHH}. The predictive model that is used by the selection mechanism of the \acs{BHH} is presented in the following section.

@@

This section presents the predictive model as implemented by the selection mechanism of the \acs{BHH}. The predictive model is arguably the most important component of the \acs{BHH}, as it drives the heuristic selection process. The \acs{BHH} implements a predictive model that predicts which heuristic to select, given the conditionality that a particular entity is used and a particular credit assignment criteria is met. The predictive model can be derived as follows.\\\\
\textbf{Setup}:
\\
Let

@@

Vector notation is dropped for sake of convenience, simplicity and readability. The parameterised predictive model, as derived from \index{Bayes' Theorem}Bayes' theorem, is then given in Equation~\eqref{eq:bhh:selection_mechanism:predictive_model} below.

@@

Notice the joint probability over events $E$ and $C$, given the selection of a heuristic $H$, denoted by the product of the sums of the separate parts for $E$ and $C$, in the denominator. The calculation in the denominator can be intractable. Since the \acs{BHH} selection mechanism uses the predictive model as a classifier, the posterior distribution as given in Equation~\eqref{eq:bhh:selection_mechanism:bayes_theorem}, can be evaluated for proportionality, as is given in Equation~\eqref{eq:bhh:selection_mechanism:predictive_model_prop_to} below.

@@

The predictive model thus models the \textit{proportional} probability of the event (selection of) heuristic $H$, given allocation to entity $E$ and credit $C$, parameterised by sampled $\theta$, $\phi$ and $\psi$. These parameters are in turn parameterised by concentrations $\alpha$, $\beta$, $\gamma_{1}$ and $\gamma_{0}$, which denote the prior beliefs of the \acs{BHH}.

@@

In Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:h_pmf} above, $N_{k}$ is a summary variable such that $N_{k} = \sum_{i=i}^{I} \mathbbm{1}_{1}(h_{i,k})$, denoting the count of occurrences of the event $h_{i}$ taking on class $k$ in $I$ independent, identical runs.

@@

In Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:EgH_pmf} above, $N_{j,k}$ is a summary variable such that $N_{j,k} = \sum_{i=i}^{I} \mathbbm{1}_{1}(e_{i,j,k})\mathbbm{1}_{1}(h_{i,k})$, denoting the count of occurrences of the events $e_{i}$ taking on class $j$ and $h_{i}$ taking on class $k$, i.e. the count of occurrences of both entity $e$ and heuristic $k$ occurring together in $I$ independent, identical runs.

@@

In Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:CgH_pmf} above, $N_{k}$ is the same summary variable as described for Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:h_pmf}. $N_{1,k}$ is a summary variable such that $N_{1,k} = \sum_{i=1}^{I} \mathbbm{1}_{1}(c_{i,k})\mathbbm{1}_{1}(h_{i,k})$, denoting the count of occurrences of the events $c_{i}$ taking on a success $(c_{i}=1)$ and $h_{i}$ taking on class $k$, i.e. the count of occurrences of both succeeding in the performance criteria and heuristic $k$ occurring together in $I$ independent, identical runs. Similarly $N_{0,k} = N_{k} - N_{1,k}$ would denote the count of occurrences of the events $c_{i}$ taking on a failure $(c_{i}=0)$ and $h_{i}$
taking on class $k$.

@@

Consider the practical implementation of the predictive model as shown in Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:HgEC_pmf} above. Computationally the equation presented in Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:HgEC_pmf} will underflow on a real computer if the resulting probabilities are very small. Discussions follow on numerical stability as it relates to practical implementation.

@@

The log-sum-exp trick as shown above caters for very small probabilities, however, there might be a situation where  a random event is never seen, purely by chance. This results in a scenario referred to as \textit{mode collapse} and is discussed next.

@@

Another suggestion is to ensure that at least one of every heuristic is always selected during training. The \acs{BHH} does not incorporate this approach as this could result in a situation where a bad choice of heuristic is continually used throughout the training process, possibly resulting in bad update steps/selections. Instead, the \acs{BHH} relies on the underlying learning process to control the selective pressure according to performance and nothing else. If this leads to a situation where mode collapse occurs, the \acs{BHH} accepts the outcome.

@@

The derivations of the \acs{MLE} for $\hat{\theta}_{k}$, $\hat{\phi}_{j,k}$ and
$\hat{\psi}_{k}$ are presented as follows. The log likelihood of $\hat{\theta}$, $\hat{\phi}$ and $\hat{\psi}$ as derived from the Equation~\eqref{eq:bhh:selection_mechanism:predictive_model_prop_to} is given in Equation~\eqref{eq:bhh:optimisation_step:mle:log_likelihood_all} below.

@@

Equation~\eqref{eq:bhh:optimisation_step:mle:log_likelihood_all} can be broken down into each of its components. Consider the log likelihood of $\psi$ as denoted by Equation~\eqref{eq:bhh:optimisation_step:mle:log_likelihood_psi} below.

@@

The \acs{MLE} for $\hat{\psi_{k}} $ can be calculated by taking the partial derivative of Equation~\eqref{eq:bhh:optimisation_step:mle:log_likelihood_psi} with respect to $\psi_{k}$ and equating to zero as presented in Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_psi} as follows.

@@

The \acs{MLE} for $\hat{\theta_{k}} $ can be calculated similarly. However, one has to compensate for the $K-1$ simplex $S$ by adding error factor $\epsilon$, to correct values for $\theta$, where $\sum_{k}^{K} \theta_{k} \neq 1$. The new log likelihood function is then given in Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_theta_part_1} below.

@@

Solving first for $\epsilon$ yields Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_theta_part_2} below.

@@

Then solving for $\theta_{k}$ yields Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_theta_part_3} as presented below.

@@

Substituting $N = \epsilon$ back into Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_theta_part_3} yields Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_theta_part_4} below.

@@

The \acs{MLE} for $\hat{\phi}_{j, k} $ is calculated similarly. To compensate for the $K-1$ simplex $S$, an error $\lambda = (\lambda_{1}, \dots, \lambda_{J})$ is added as is shown in Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_phi_part_1} below.

@@

Solving first for $\lambda_{j}$ yields Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_phi_part_2} as is given below.

@@

Then solving for $\phi_{j,k}$ yields Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_phi_part_3} below.

@@

Substituting $N_{j} = \lambda_{j}$ back into Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_phi_part_3} yields Equation~\eqref{eq:bhh:optimisation_step:mle:proof_log_likelihood_phi_part_4} below.

@@

The event $H$ is a multinomial distribution with probability parameter $\theta$. It is known that the conjugate prior to a multinomial distribution is a \index{Dirichlet probability distribution}Dirichlet probability distribution. The \textit{prior} probability distribution for $\theta_{k}$ is thus presented in Equation~\eqref{eq:bhh:optimisation_step:map:map_theta_prior} below.

@@

Furthermore, the event $E$ is also a multinomial distribution with parameter $\phi$. The \textit{prior} probability distribution for $\phi_{j,k}$ is thus presented in Equation~\eqref{eq:bhh:optimisation_step:map:map_phi_prior} below.

@@

Finally, the event $C$ is a \index{Binomial probability distribution}Binomial probability distribution with probability parameter $\psi$. It is known that the conjugate prior to a \index{Binomial probability distribution}Binomial probability distribution is a Beta probability distribution. The prior probability distribution for $\psi_{k}$ is thus presented in Equation~\eqref{eq:bhh:optimisation_step:map:map_psi_prior} below.

@@

Putting the likelihood and priors together yields the posterior distribution as given in Equation~\eqref{eq:bhh:optimisation_step:map:map_psi_prior} below.

@@

From Equation~\eqref{eq:bhh:optimisation_step:map:posterior} above, it can be seen that the posterior distribution has the same form $\mathcal{A}(v)$ as the prior distribution. The concentration update operations can then be given in Equations \eqref{eq:bhh:optimisation_step:map:alpha_update_operation}, \eqref{eq:bhh:optimisation_step:map:beta_update_operation}, \eqref{eq:bhh:optimisation_step:map:gamma1_update_operation} and \eqref{eq:bhh:optimisation_step:map:gamma2_update_operation} below.

@@

It can be seen that the prior and posterior probability distributions are of the same form, $\mathcal{A}(v)$. It can therefore be said that the \acs{BHH} implements a Gaussian process. Since the reselection of heuristics happens at regular intervals, the outcome of a selection in one iteration may influence the outcome of another in the next iteration, making the implementation of the \acs{BHH} a \acf{HMM}.

@@

The design choices at a high level is not as strict for the \acs{BHH} as it is for low-level heuristics. If one is unsure as to which low-level hyper-parameters to use, one could simply include multiple configurations of a low-level heuristic, each with its own unique hyper-parameter configurations.

@@

A number of other hyper-parameters are specifically included to address the list of considerations as given above. These include \textit{reselection}, \textit{replay} and \textit{reanalysis window sizes}, \textit{burn in}, \textit{discounted rewards} and \textit{normalisation}. Each of these are presented next.

@@

The following section discusses the default values for proxied update step operations.
